### 2020-04-15 probabilisticProgrammingMartingales.tex
Some comments and questions.

\lo{Your operational semantics does not maintain a record of the current weight of the reduction.
Note that a.s.~termination does depend on $\tscore$: see \cite[\S 4.3]{DBLP:journals/corr/abs-2004-03924}\footnote{\url{https://arxiv.org/abs/2004.03924}}.}

@@

@@

Given $M'_n$, there is some finite bound on the number of reduction steps that can take place from $M'_n$ without a $\tY$-reduction step, because of the type system, therefore $y_{n+1}$ is (conditional on $\mathcal{F}_{y_n+1}$) a bounded stopping time, therefore $\mathbb{E}[f(M_{y_{n+1}}) \mid \mathcal{F}_{y_n+1}] \leq f(M_{y_n+1})$. If $M_{y_n}$ isn't already in normal form, then $M_{y_n} = E[\tY N]$ for some $E, N$, therefore $M_{y_n+1} = N (\tY N)$ and $f(M_{y_n+1}) \leq f(M_{y_n}) - 1$, therefore if $M_{y_n}$ isn't normal form, $\mathbb{E}[X_{n+1} \mid \mathcal{F}_{y_n+1}] \leq X_n - 1$.

Overall, this implies that $\mathbb{E}[X_{n+1}] - \mathbb{E}[X_{n}] \leq -\mathbb P[M_{y_n} \text{not n.f.}]$ therefore as $\mathbb{E}[X_n]$ is bounded below, $\mathbb P[M_{y_n} \text{not n.f.}]$ must tend to 0 as $n \to \infty$. $M_{y_n} \text{n.f.} \Rightarrow M_{y_{n+1}} \text{n.f.}$ therefore 
\begin{align*}
\mathbb P[M_{y_n} \text{not n.f.~for infinitely many values of $n$}]  
& = \mathbb P[\forall n: M_{y_n} \text{not n.f.}] \\
& \leq \inf_n \mathbb P[M_{y_n} \text{not n.f.}] \\
& = 0
\end{align*} 
therefore $M$ terminates almost surely.
\end{proof}

>>>>

\lo{Your operational semantics does not maintain a record of the current weight of the reduction.
Note that a.s.~termination does depend on $\tscore$: see \cite[\S 4.3]{DBLP:journals/corr/abs-2004-03924}\footnote{\url{https://arxiv.org/abs/2004.03924}}.
I think it important to take the behaviour of $\tscore$ into account, perhaps as a future task.}

@@

\lo{By ``normal form'' in the side condition above, you mean value? (Normal form can be confusing.)}

@@

Given $M'_n$, there is some finite bound on the number of reduction steps that can take place from $M'_n$ without a $\tY$-reduction step, because of the type system \lo{Why?}, therefore $y_{n+1}$ is (conditional on $\mathcal{F}_{y_n+1}$) a bounded stopping time, therefore $\mathbb{E}[f(M_{y_{n+1}}) \mid \mathcal{F}_{y_n+1}] \leq f(M_{y_n+1})$. If $M_{y_n}$ isn't already in normal form, then $M_{y_n} = E[\tY N]$ for some $E, N$, therefore $M_{y_n+1} = N (\tY N)$ and $f(M_{y_n+1}) \leq f(M_{y_n}) - 1$, therefore if $M_{y_n}$ isn't normal form, $\mathbb{E}[X_{n+1} \mid \mathcal{F}_{y_n+1}] \leq X_n - 1$.

Overall, this implies \lo{Why?} that $\mathbb{E}[X_{n+1}] - \mathbb{E}[X_{n}] \leq -\mathbb P[M_{y_n} \text{not n.f.}]$ therefore as $\mathbb{E}[X_n]$ is bounded below, $\mathbb P[M_{y_n} \text{not n.f.}]$ must tend to 0 as $n \to \infty$. $M_{y_n} \text{n.f.} \Rightarrow M_{y_{n+1}} \text{n.f.}$ therefore 
\begin{align*}
\mathbb P[M_{y_n} \text{not n.f.~for infinitely many values of $n$}]  
& = \mathbb P[\forall n: M_{y_n} \text{not n.f.}] \\
& \leq \inf_n \mathbb P[M_{y_n} \text{not n.f.}] \\
& = 0
\end{align*} 
therefore $M$ terminates almost surely.
\end{proof}
